---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%"
)
```

# MRCO
**M**ulti **R**esolution **C**luster **O**ptimization
<!-- badges: start -->
<!-- badges: end -->


Visualize your clustering of multiple independently generated resolutions (or sensitivities) in a stacked tree like graph.
It shows if clusters remain in similar order across resolution change of reiterations using the same clustering algorithm.
Color clusters as pie-charts based on their sample proportions, if sample level metadata is included in the input, to see if cluster splitting is correlating with a split of metadata groups.
If samples remain in same cluster grouping across many resolution 'steps', we assume a cluster to be stable.
An automated cluster selection algorithm is implemented to select the 'most stable' clusters relative within the given data.
Samples that frequently change between clusters upon change in resolution depict an uncertain state.
Therefore, a 'debris' cluster is created if samples are not part of any of the selected stable clusters.   
The software was tested on single cell RNA-Sequencing data, but can be applied to any type of data (on your own risk).
The general concept remains the same: A clustering algorithm can cluster your data similar multiple times, the similarity between multiple reiterations can be used to infer the most stable configuration for your experimental scope.

# Installation

You can install the development version of MRCO from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("peterwolf4/MRCO")
```

# Example


```{r example data}

# Create Example dataframe

# Required:
# Dataframe or Matrix of samples (rows) and increasing clustering resolutions (columns).
# In single cell this will be: cells (rows) and increasing clustering resolution (columns)

example_clustering <- data.frame(row.names = paste0("cell_",seq(1,10,1)),
                                  check.names = F , "0.1" = rep(0,10), 
                                  "0.2" = c(rep(1,4), rep(0,6)), 
                                  "0.3" = c(rep(0,4),rep(1,3), rep(2,3)),
                                  "0.4" = c(3,rep(0,3), rep(1,3),2,2,3))

# Optional:
# Extra metadata variables (columns) for your samples (rows). 
# Variable can be continuous (automatically binned) or discrete. 

# Setting the seed for reproducable sample results in this example.

set.seed(1)

# Creating example 
example_metadata <- data.frame(
                       "quality" = sample(c("high","medium","low"),10, T),
                       "batch" = sample(c("1","2","3"),10, T),
                       "sample" = sample(paste("cell",seq(1,10), sep = "_"), 10),
                       "expr_A" = sample(seq(100,1000), 10, T),
                       "expr_B" = sample(seq(1,10), 10, T))

# Joining cluster results with metadata
# **Be sure the order of your samples align**, in this example, they do not.

head(example_clustering, 3)
head(example_metadata, 3)

# To be sure we can order both dataframes.
# Single cell clustering- and meta-data should be ordered by default already.

example_data <- cbind(example_metadata[order(example_metadata$sample),],
                      example_clustering[order(rownames(example_clustering)),])
# Update rownames, numbering is obscurred in this example by the order function
rownames(example_data) <- seq_len(nrow(example_data))
head(example_data, 3)
```


```{r A basic MRCO run}
# Load MRCO
library(MRCO)

# Running MRCO
# We can run the function with just the clustering results:
# Lowest resolution clustering has one cluster and the highest resolution 4.
# This example could be the result of 4 k-mers clusterings with k = 1,2,3,4.
MRCO(example_clustering)

# MRCO is verbose when it comes to warnings. We can also use our merged data. 
# Missing inputs prefix and/or suffix to target clustering result columns.
MRCO(example_data, prefix = "0.")

# We can plot metadata variables as piecharts and see how samples cluster.
MRCO(example_data, metadata_column_name = "quality",prefix = "0.")
# The example data shows no clear seperation based on sample quality, good sign.

MRCO(example_data, metadata_column_name = "expr_A",prefix = "0.")
# Example A: graph hints that cluster 3 at resolution step 4 (4_3);
# is split based on expression values of A. (All samples in 4_3 have high values of A)

MRCO(example_data, metadata_column_name = "expr_B",prefix = "0.")
# Example B: graph hints that lefthand clusters have rather high-;
# while lefthand have rather low- values of B.

# Given examples are of course no final conclusions.
# Rather, they are for a quick visual overview of your variables of interest, 
# and if they correlate to your clustering results.


# Selecting clusters of interest for further downstream analysis:
# Imagine we want to retain all samples of 4_3 as oner cluster to investigate A,
# but also need all other samples in second step clusters before the separation of B.
# MRCO enables you to pick any of the clusters manually at wish.
# Use merge_downwards (default) to preserve clusters from lower resolutions.
MRCO_clustering <- MRCO(example_data, prefix = "0.", 
     nodes_selection = c("4_3","2_0","2_1"),
     merge_downwards = FALSE)

# MRCO highlights selected clusters visually. 
# Each sample now has a unique cluster number composed of 'resolution'_'cluster'.
# Output is returned in order of input.
head(MRCO_clustering$nodes_selected$cells_NonGlobalClustering,3)
table(MRCO_clustering$nodes_selected$cells_NonGlobalClustering$id)

# Automated Stable Cluster Detection
# Note that stable clusters are relative to the data and should still be checked!
# Note that this example data is too small for the algorithm to be truthfully useful.
MRCO_clustering <- MRCO(example_data, prefix = "0.", suggest_cut = TRUE)

# Stable clusters are representative for the trees 'branches'. 
# We can check which clusters were selected as stable, just as before:
MRCO_clustering$nodes_selected$nodes_selection
# We see that automated suggestion found the first cluster to be 'stable' already.
# This is not a problem though, since merge_downwards = FALSE prefers the sub clusters.
head(MRCO_clustering$nodes_selected$cells_NonGlobalClustering, 3)

# In larger datasets you likely won't see general clusters selected as stable,
# since more information is available to detect the most stable candidates.

```

## Disclaimer
Package is currently in development. Planned for release on bioconductor. 
Good & reliable single cell clustering results have already been found. 
Do use at your own caution though! Feel free to reach out regarding bugs and ideas.


